---
title: "news"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{news}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gcamfaostat)
```

# Goals for the next version include:
2023-09-25  

* Improve the documentation and the visualization of the *module_aglu_L100.FAO_SUA_PrimaryEquivalent* module
  + The module developed a new method of primary commodity equivalent aggregation to aggregate supply-utilization-accounting (SUA) data for items along the supply chain (e.g., wheat flour, bran, and germ to wheat-primary-equivalent). 
  + The method preserves balance across space (trade balance), time (storage carryover), supply-utilization, and the combination of these dimensions with minimal adjustments. 
  + We have applies the new method of primary equivalent aggregation to aggregating FAO ~500 SUA (SCL) items to ~100 primary equivalent items in FAO Food Balance Sheet (FBS).
* Compare the results from the new primary commodity equivalent method with the FBS data compiled using FAOSTAT method (never disclosed).


# gcamfaostat 1.0.0

**First release**
The first release of gcamfaostat 1.0.0 includes the data generated for the Global Change Analysis Model v7.0 [GCAM v7.0](https://github.com/JGCRI/gcam-core/releases/tag/gcam-v7.0). 

**Raw data archive**
2023-08-19  

The source data downloaded from FAOSTAT is archived at a [Zenodo repository]( https://zenodo.org/deposit/8260225).

