---
title: "Preparing data for gcamfaostat"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: references.bib  
vignette: >
  %\VignetteIndexEntry{Preparing data for gcamfaostat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

The `gcamfaostat` tool is designed to streamline the processing and synthesis of raw data sourced from [FAOSTAT](https://www.fao.org/faostat/en/#data). The initial phase of this process involves data procurement, with a critical awareness of FAOSTAT's ongoing data updates. These updates encompass a spectrum of improvements, including, e.g., the addition of historical data for new countries and items, data completion for previously missing entries, and refinements such as changes in country nomenclature. This dynamic data landscape underscores the need for a robust and adaptable approach in the `gcamfaostat` workflow. Here we describe a few key functions in `gcamdatafaostat` created to procure the raw data and facilitate the processing.

When `gcamfaostat` is downloaded, preprocessed FAOSTAT data, i.e., output of the `xfaostat_L101_RawDataPreProc*` modules, are stored in the [Prebuilt Data](https://github.com/JGCRI/gcamfaostat/blob/main/data/PREBUILT_DATA.rda) of the package. 

* The package can be run with using those data, but make sure `Process_Raw_FAO_Data <- FALSE` in `constants.R`.  
* The prebuilt data were generated using FAOSTAT data archived in Zenodo address specified in the `FF_download_RemoteArchive` function [@xin_zhao_2023_8260225].


### 1. Get the latest FAOSTAT Metadata  
[`FAOSTAT_metadata()`](https://jgcri.github.io/gcamfaostat/reference/FAOSTAT_metadata.html)  

* Accessing API and returning a data frame of metadata

### 2. Generate the metadata for the `gcamfaostat` input data  

[`gcamfaostat_metadata()`](https://jgcri.github.io/gcamfaostat/reference/gcamfaostat_metadata.html)  

* The function will save the latest FAOSTAT metadata to the [metadata_log](https://github.com/JGCRI/gcamfaostat/tree/main/inst/extdata/FAOSTAT/metadata_log)  
* The function will access both the latest FAOSTAT metadata and local data information and returns a summary table including the dataset information needed for gcamfaostat (see **Table 1** below).  
* The dataset code needed were specified in the function to get a subset of the FAOSTAT metadata. The function will return only dataset code required when setting `OnlyReturnDatasetCodeRequired = FALSE`.  
* The function will check whether FAOSTAT raw data exists locally (`Exist_Local`) and in [Prebuilt Data](https://github.com/JGCRI/gcamfaostat/blob/main/data/PREBUILT_DATA.rda) (`Exist_Prebuilt`). If `Exist_Prebuilt` is `TRUE` for all dataset, the package is ready to be built based on the Prebuilt package data.   
* `FAO update data` and `FAO size` indicate the information based on the latest FAOSTAT metadata.  

  
**Table 1. FAOSTAT dataset processed in gcamfaostat**
``` {r, echo=FALSE, warning=FALSE, message=FALSE}
# devtools::load_all()
library(gcamfaostat)
library(dplyr)
library(kableExtra)
gcamfaostat_metadata(.DIR_RAW_DATA_FAOSTAT = file.path("..", "inst/extdata/FAOSTAT")) %>% 
  select(-`Local file size`) -> p

p %>%
  #kbl(caption = '<b>FAOSTAT dataset processed in gcamfaostat</b>', format = 'html') %>%
  kbl(format = 'html') %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

Note that if Exist_Prebuilt is `TRUE` for a dataset, it suggests the raw data was saved in the Prebuilt data. And if Exist_Prebuilt is `TRUE` for all dataset, the package is ready to be built based on the Prebuilt data.


### 3. Download FAOSTAT raw data  

[`FF_download_RemoteArchive()`](https://jgcri.github.io/gcamfaostat/reference/FF_download_RemoteArchive.html)  

* The function downloads the FAOSTAT raw data needed for the package from a remote archive.   
* The default Zenodo archive currently included in the function includes a snapshot of FAOSTAT data to ensure replicability.   
*  The archived data is consistent with the Prebuilt package data.

[`FF_download_FAOSTAT()`](https://jgcri.github.io/gcamfaostat/reference/FF_download_FAOSTAT.html)
* The function downloads the latest raw data from FAOSTAT.
 
Two functions above are created for downloading the raw data from a remote archive or the FAOSTAT API (latest data). The dataset code variable in the two functions, if including all, can be generated using `gcamfaostat_metadata(OnlyReturnDatasetCodeRequired = T)`. 

Example:
``` {r, eval =FALSE}
# Dataset PP, producer prices, is downloaded from RemoteArchiveURL to DATA_FOLDER
# RemoteArchiveURL = "https://zenodo.org/record/13941470/files/"
# DATA_FOLDER = "inst/extdata/FAOSTAT"
FF_download_RemoteArchive( 
  DATASETCODE = "PP",   
  OverWrite = TRUE     # overwrite existing PP dataset 
  )

# Dataset OA, population, is downloaded from FAOSTAT to DATA_FOLDER
FF_download_FAOSTAT(DATASETCODE = "OA", OverWrite = TRUE)

# Note that single DATASETCODE is allowed in both functions. 
```

### 4. Check local raw data info

[`FF_rawdata_info()`](https://jgcri.github.io/gcamfaostat/reference/FF_rawdata_info.html)  

* Providing more detailed metadata information, similar to `gcamfaostat_metadata` (which calls `FF_rawdata_info`).  
* Indicate whether an update from FAOSTAT is potentially needed.
* Can be used to download data if not exist.

Example:
``` {r, eval =FALSE}
# Provide detailed metadata of "PP" and "OA" in DATA_FOLDER ("inst/extdata/FAOSTAT")
FF_rawdata_info(DATASETCODE = c("PP", "OA"), DOWNLOAD_NONEXIST = FALSE)

# If "PP" or "OA" does not exist, download from remote archive
FF_rawdata_info(DATASETCODE = c("PP", "OA"), DOWNLOAD_NONEXIST = TRUE, FAOSTAT_or_Archive = "Archive")
```

### 5. Load raw data into package  

[`FAOSTAT_load_raw_data`](https://jgcri.github.io/gcamfaostat/reference/FAOSTAT_load_raw_data.html)  

* Loading FAOSTAT raw data, e.g., used in `xfaostat_L191_RawDataPreProc*` modules. 
* Note that `.` in column name is substituted with `_`.

Example:
``` {r, eval =FALSE}
# Read raw data of "PP" and "OA" from DATA_FOLDER ("inst/extdata/FAOSTAT") to .GlobalEnv
FAOSTAT_load_raw_data(DATASETCODE = c("PP", "OA"), .Envir = .GlobalEnv)
```

### 6. Update the Prebulit data

If users updated the FAOSTAT raw data, and run `driver_drake` is run with `Process_Raw_FAO_Data <- TRUE` in `constants.R`, the updated preprecessed data are stored in drake cache. And they can be used to update `Prebulit data` by sourcing `data-raw/generate_package_data.R`.

## References
